\LARGE \bf Mean Estimation from Single-bit Measurements

### 2017-09-26 Writeups/onebit_distributed_V3.tex
Add files via upload

% Lower Bound on the assymptotic minimal effeciency.
% Scheme that attains the assymptotic minimal effeciency. 
% Bayesian scheme that is one step optimal. We leave open the question whether this scheme attains the asymptotic minimal effeciency

\section{Adaptive Estimation \label{sec:sequential}}

The first main results of this paper, as described in Theorem~\ref{thm:adpative_lower_bound} below, states that the asymptotic relative efficiency of and adaptive estimator cannot get below $\pi/2$. Next, we provide a particular adaptive estimation scheme and show in Theorem~\ref{thm:sgd} that asymptotically attains this efficiency with equality. Finally, in Theorem~\ref{thm:opt_one_step} we provide an adaptive estimation scheme that is optimal in the sense that at each step $i$, the message $M_i$ that minimizes the MSE given $X_i$ and the previous $M^{i-1}$ messages is chosen. 
%While we provide expressions for the optimal message, evaluating the asymptotic efficiency of this estimation scheme turns out to be a challenging task. We therefore leave open the question whether this scheme attains the lower bound with equality, and only provide numerical result that support this conjecture.  \\

%\begin{figure}
%\begin{center}
%\begin{tikzpicture}[node distance=2cm,auto,>=latex]
%  \node at (0,0) (theta) {$\theta$} ;
%  \node at (2,0) (x1) {$X_1$};
%  \node[below of = x1, node distance = 1cm] (x2) {$X_2$};
%    \node[below of = x2, node distance = 1cm] (dots) {$\vdots$};
%  \node[below of = dots, node distance = 1cm] (xn) {$X_n$};
%  
%   \node[right of = x1, node distance = 2cm] (m1) {$M_1$};
%  \node[right of = x2, node distance = 2cm] (m2) {$M_2$};
%    \node[right of = xn, node distance = 2cm] (mn) {$M_n$};
%
%\draw[->] (theta) -- (x1);
%\draw[->] (theta) -- (x2);
%\draw[->] (theta) -- (xn);
%
%\draw[->] (x1)--(m1);
%\draw[->] (x2)--(m2);
%\draw[->] (xn)--(mn);
%
%\draw[->] (m1)--(m2);
%\draw[->, dashed] (m2)--(mn); 
%  
%\end{tikzpicture}
%\end{center}
%\caption{\label{fig:sequencial_graphical} Graphical model notation of the sequential setting.}
%\end{figure}

%\subsection{Properties of an Optimal Scheme}
%The sequential nature of the encoding and the estimation is best explain by the notions of actions and rewards taken from dynamic programming. The \emph{action} taken by the $i$th encoder is the design of a detection region $M_i^{-1}(1)$ that would lead to $M_i = 1$ assuming when $X_i \in M_i^{-1}(1)$ or $M_i=-1$ otherwise. The \emph{reward} (or penalty) is only evaluated at the end of the time horizon $n$.
%% Dynamic program. State = $M^n$. 
%% Reward is collected at the end
%%

\subsection{A Lower Bound on Adaptive Single-Bit Schemes}
Our first results asserts that the relative efficiency \eqref{eq:relative_efficiency} of any adaptive estimation scheme is bounded from below by $\pi/2$, as follows from the following theorem:
\begin{thm}[minimal relative effeciency] \label{thm:adpative_lower_bound}
Let $\widehat{\theta}_n$ be any estimator of $\theta$ in the adaptive setting of Fig.~\ref{fig:sequential}.
% Let $\pi(\theta)$ be an absolutely continuous prior distribution on $\Theta$ with density function converges to zero at the endpoints of the interval $\Theta$. 
Assumes that $P(d\theta)$ converges to zero at the endpoints of the interval $\Theta$. Then
\[
\mathbb E (\theta-\theta_n)^2 \geq  \frac{\pi \sigma^2 }{2n +\pi \sigma^2  I_0}  =  \frac{\pi}{2n}\sigma^2+o(n^{-1}),
\]
where 
\[
I_0 = \mathbb E \left( \frac{d}{d\theta} \log P_\theta(\theta) \right)^2
\]
is the Fisher information with respect to a location model in $\theta$. 
\end{thm}

\subsubsection*{Sketch of Proof}
The main idea in the proof is to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$. Once this bound is achieved, the result follows by using the Van-Trees inequality \cite{gill1995applications} which bounds from below the MSE of any estimator of $\theta$ by the inverse of the expected value of the aforementioned Fisher information plus $I_0$. The details are given in the Appendix.\\

Next, we present an adaptive estimation scheme which attains the minimal asymptotic efficiency $\pi/2$. 
%Theorem~\ref{thm:adpative_lower_bound}. This estimation scheme can be seen as a gradient descent

\subsection{Asymptotically Optimal Estimator}
Consider the following estimator $\widehat{\theta}_n$ for $\theta$:  set 
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n \sgn (X_n - \theta_n), \quad n = 1,2,\ldots,
\end{equation}
where $\left\{\gamma_n \right\}_{n=1}^\infty$ is any strictly positive sequence satisfying 
\begin{enumerate}
\item[(i)] $\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n)$ \\
\item[(ii)] $\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty$ 
for some $0< \lambda \leq 1$.
\end{enumerate}
(e.g. $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). The $n$th step estimation is defined by 
\begin{equation} \label{eq:sgd_est}
\widehat{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

As the following theorem shows, the estimator $\widehat{\theta}_n$ defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg} attains the minimal asymptotic efficiency as established by Theorem~\ref{thm:adpative_lower_bound}.
\begin{thm} \label{thm:sgd}
The sequence $\widehat{\theta}_n$ of \eqref{eq:sgd_alg} satisfies
\[
\sqrt{n} \left( \widehat{\theta}_n - \theta \right) \overset{d}{\rightarrow} \mathcal N \left(0,  \pi \sigma^2 /2 \right).
\]
\end{thm}

\subsubsection*{Proof}
The asymptotic behavior of \eqref{eq:sgd_est} is a special case of  \cite[Thm. 4]{polyak1992acceleration}. The details are provided in the Appendix.\\

Note that $\theta_0$ is not explicitly defined in 
in equation \eqref{eq:sgd_est}. Arguably, a reasonable initialization is $\theta_0 = \mathbb E [\theta]$. However, Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator, is indifferent to this initialization. Thus, the optimal efficiency is attained without using the prior distribution on $\theta$. Nevertheless, the bound in Theorem~\ref{thm:adpative_lower_bound} suggests that the estimation error can be greatly reduced whenever the prior of $\theta$ provides a lot of information $I_0$ on its location. In order to incorporate this information in a more meaningful way, we present a Bayesian estimation scheme that is conjectured to attain the optimal asymptotic efficiency.

\subsection{One-step Optimal Estimation}
In this subsection we present an estimation scheme that posses the property of \emph{one-step optimality}. That is, at each step $i$, the $i$th encoder designs the detection region $M_i^{-1}(1)$ such that the MSE given $M^i$ is minimal. In other word, this scheme designs the messages in a greedy manner, such that the MSE at step $i$ is minimal given the current state of the estimation described by $M^{i-1}$. \\

The structure of the message that minimizes the next step MSE is given by the following theorem:
\begin{thm}[optimal one-step estimation] \label{thm:opt_one_step}
Let $\pi(\theta)$ be an absolutely continuous log-concave probability distribution. Given a sample $X$ from the distribution $\mathcal N(\theta, \sigma^2)$, define 
\begin{equation}
\label{eq:adaptive_main_message}
M = \sgn(X - \tau),
\end{equation}
where $\tau$ satisfies the equation
\begin{equation}
 \label{eq:fixed_point}
 \tau = \frac{m^-(\tau) + m^+(\tau)}{2},
\end{equation}
with
\begin{align*}
m^-(\tau)  & = \frac{\int_{-\infty}^{\tau} \theta P(d\theta) }{\int_{-\infty}^{\tau} P(d\theta)} ,\\
m^+(\tau) & = \frac{\int_{\tau}^\infty \theta P(d\theta) }{\int_{\tau}^\infty P(d\theta)} .
\end{align*}
Then for any estimator $\widehat{\theta}$ which is a function of $M'(X) \in \{-1,1\}$, we have
\begin{equation}
\label{eq:opt_cond}
\mathbb E \left(\theta-\widehat{\theta}(M')\right)^2 \geq  \mathbb E \left(\theta- \mathbb E[\theta|M]\right)^2,
\end{equation}
\end{thm}

\begin{proof}
The proof is completed by the following two lemmas, proofs of which can be found in the Appendix:
\begin{lem} \label{lem:unique}
Let $f(x)$ be a log-concave probability density function. Then the equation 
\begin{equation}
\label{eq:lem_fixed_point}
2x = \frac{\int_x^\infty uf(u)du}{\int_x^\infty f(u)du} + \frac{\int_{-\infty}^x uf(u)du}{\int_{-\infty}^x f(u)du} 
\end{equation}
has a unique solution.
\end{lem}
\end{proof}

\begin{lem} \label{lem:adaptive}
Let $U$ be an absolutely continuous random variable with
pdf $P(du)$. Then the one-bit message $M^\star$ that minimizes
\[
\int \left( u - \mathbb E[U|M(u)]  \right)^2 P(du)
\]
is given by
\[
M^\star  =  \sgn(U - \tau),
\]
where $\tau$ is the unique solution to the equation
 \[
2 \tau = \frac{\int_{\tau}^\infty u P(du)} {\int_{\tau}^\infty P(du)} + \frac{\int_{-\infty}^{\tau} u P(du)}{\int_{-\infty}^{\tau} P(du)}.
\]
\end{lem}

Theorem~\ref{thm:opt_one_step} suggests the following Bayesian estimation scheme: set $P_0(t) = \pi(\theta)$, and for $n\geq 1$ update the prior as
\begin{align}
P_n(t) = & P(\theta=t |M^n) = \frac{ P\left( \theta=t | M^{n-1} \right) P(M_n | \theta = t , M^{n-1})  } { P(M_n | M^{n-1} )} \nonumber \\ 
& = \alpha_n  P_{n-1}(t) \Phi\left(M_n \frac{ t - \tau_{n-1} }{\sigma} \right), \label{eq:density_update}
\end{align}
where $\alpha_n$ is a normalization coefficient that equals to
\[
\alpha_n = \left(\int_{\mathbb R} P_{n-1}(t) \Phi\left(M_n \frac{t- \tau_{n-1} }{\sigma} \right)  dt \right)^{-1}. 
\]
The $n$th estimate for $\theta$ is the conditional expectation of $\theta$ given $M^n$, namely
\begin{equation}
\theta_n = \mathbb E \left[ \theta| M^n\right] = \int_{-\infty}^\infty t P_n(t) dt. \label{eq:estimator_update}
\end{equation}
Next, solve equation \eqref{eq:fixed_point} with the updated prior $P_n(t)$ instead of $P(\theta)$. Note that since the standard normal cdf $\Phi(x)$ is log-concave, the updated prior $P_n(t)$ remains log-concave and thus a unique solution to \eqref{eq:fixed_point} is guaranteed by Lemma~\ref{lem:unique}. Finally, update the $n+1$th message as
\begin{equation}\label{eq:message_update}
M_{n+1} = \sgn(X_{n+1}-\tau_n).
\end{equation}

Since the equation \eqref{eq:fixed_point} has no analytic solution in general, it is hard to derive the asymptotic behavior of the Bayes estimator defined by \eqref{eq:estimator_update} and \eqref{eq:message_update}. We conjecture, however, that it attains the asymptotic relative efficiency of $\sigma^2\pi/2$, as can be observed from the numerical simulation illustrated in Fig.~\ref{fig:adaptive_error}. Also shown in Fig.~\ref{fig:adaptive_error} are the normalized MSE of the estimator defined by \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, as well as the MSE achieved by the sample mean for the same sample realization.

 % In fact, the numerical simulation in Fig.~\ref{fig:simulation} shows that this scheme outperform the estimator of \eqref{eq:sgd_est} even when the prior on $\Theta$ is least favorable for Bayes estimation \cite{Bickel1981}.

\begin{figure}
\begin{center}
\begin{tikzpicture}
\node at (0,0) {\includegraphics[scale=0.4]{error}};
\end{tikzpicture}
\caption{Normalized empirical risk $n\left(\widehat{\theta}_n-\theta\right)^2$ versus number of samples $n$ for $100$ Monte Carlo trials, where $\theta$ is chosen uniformly in the interval $(-5,5)$. \label{fig:adaptive_error}  }
\end{center}
\end{figure}

\begin{comment}
\subsection{Relation to Sigma-Delta Modulation \label{subsec:sdm}}
The SDM is a device that converts continuous-time analog signals to a discrete-time sequence of binary digits. The basic operation of a SDM can be described as follows: the modulator takes as input a sequence of continuous amplitude samples $X_n$, which are typically represents samples of a continuous-time signal taken uniformly at sampling rate $f_s$ much higher than the Nyquist rate of the signal \cite{1095151}. The output of the modulator is a sequence $M_n$ of binary values, obtained according to the following rule: 
\begin{align}
\begin{cases} V_{n+1} & =  V_n - \alpha q(V_n) + X_n, \\
M_n & = \sgn(V_n)  \end{cases}, \quad n=1,2,\ldots, \label{eq:sdm}
\end{align}
where $\alpha > 0$ is a parameter of the modulator that is determined by the maximal value of the input sequence $X^n$. The SDM of the form \eqref{eq:sdm} was studied in \cite{53738} under the assumption that the input process $X_n$ is an i.i.d Gaussian process. The main results from \cite{53738} shows that regardless of the initial state $V_0$, the process $M_n$ is a stationary ergodic process with mean $\theta$. It follows that a consistent estimation of $\theta$ is obtained by taking the mean of the sequence $\{M_n\}$. However, the results from \cite{53738} does not provide a closed form for the second order statistics of $\{M_n\}$, so that convergence rate cannot be derived from \cite{53738}.  
\end{comment}

@@

\subsection{Threshold Detection}
We assume that each message is of the form
\[
M_i = \sgn(t_i - X_i) = \begin{cases} 1 & X_i< t_i, \\
-1 & X_i > t_i,
\end{cases}  
\]
where $t_i\in\mathbb R$ is the \emph{threshold} of the $i$th detector applied by the encoder on it sample $X_i$. In order to characterize the asymptotic behavior of estimation from these sequence of messages, we consider the \emph{density} of $n$ threshold values defined as:
\[
\lambda_n([a,b]) = \frac{1}{n} \left| \mathcal T \cap [a,b] \right|,
\]
where $\mathcal T  = \left\{t_n, n \in \mathbb N \right\}$ is the collection of all threshold values. We further assume that $\lambda_n$ converges (weakly) to a probability measure $\lambda(t)$ on $\mathbb R$.  \par
In order to estimate $\theta$ from the messages $M^n$, we consider the maximum likelihood (ML) estimator. The log-likelihood function of $\theta$ is given by
\[
l(M^n  |\theta ) =  \sum_{i=1}^n \log \left( \Phi \left( M_i \frac{t_i - \theta}{\sigma} \right) \right). 
\]
Since $\Phi(x)$ is a log concave function, the log-likelihood function has a unique maximizer $\widehat{\theta}_n$ which is the ML estimator.  \\

@@

\subsubsection{Worst case $\theta$} in order to bound the local asymptotic minimax risk it is required to consider the density $\lambda$ that minimizes the worst case $\theta$ in the parameter space $\Theta$. We assume that $\Theta = [-b,b]$ for some $b>0$, and consider 
\begin{equation} \label{eq:variational}
\sup_{\lambda} \inf_{\theta \in [-b,b]}  {K(\theta)} = \int \eta\left( \frac{t-\theta}{\sigma} \right) \lambda (dt),
\end{equation}
where the suprimum is over all probability measure $\lambda$ on $\mathbb R$. Since \eqref{eq:variational} is concave in $\lambda$, this optimization problem can be solved using a convex program. 
As illustrated in Fig.~\ref{fig:cvx}, the threshold density that maximizes \eqref{eq:variational} is supported over a discrete set in the interval $[-b,b]$. Fig.~\ref{fig:cvx2} illustrates the asymptotic ML risk as a function of the length $2b$ of the parameter space $\theta$.

@@

As can be seen from Fig.~\ref{fig:opt_density}, for a normal prior with a small variance compared to the variance of the samples $\sigma^2$, the optimal density $\lambda$ is a mass distribution. As it turns out from the following proposition, the choice of  $\lambda$ to be a mass distribution leads to a general upper bound on the expected asymptotic risk, and not only on the minimal value of \eqref{eq:cvx_average}. 
\begin{prop}\label{prop:upper_bound}
Let $\theta_0 = \mathbb E \theta$. Then
\begin{equation}
\label{eq:upper_bound}
\mathbb E \frac{\sigma^2}{K(\theta)}  \leq \sigma^2 \int \frac{\pi(d\theta)}{\eta \left( \frac{\theta_0 - \theta}{\sigma} \right)}, 
\end{equation}
\end{prop}
\begin{proof}
Since the function $x \rightarrow 1/x$ is convex for positive values, Jensen's inequality applied to the RHS of \eqref{eq:cvx_average} leads to
\[
 \frac{\sigma^2}{K(\theta)} \leq \sigma^2 \int  \frac{ \lambda(dt)}{ \eta \left( \frac{t-\theta}{\sigma} \right)  }. 
\]
Therefore, the expected value of $\sigma^2/K(\theta)$ satisfies
\begin{align}
\mathbb E  \frac{\sigma^2}{K(\theta)}  \leq \sigma^2  \int \int \frac{\pi(d\theta) \lambda(dt) }{\eta \left( \frac{t - \theta}{\sigma} \right)}. \label{eq:upper_bound_proof}
\end{align}
The bound \eqref{eq:upper_bound} follows by using $\lambda = \delta_{\theta_0}$. 
\end{proof}
We note that since the function $1/\eta(x)$ is quasi-convex and symmetric, taking $\lambda = \lambda_{\theta_0}$ minimizes the RHS of \eqref{eq:upper_bound_proof} and leading to the tightest bound among the family bounds obtained after using Jensen's inequality.  \par
The bound \eqref{eq:upper_bound} is not trivial as long as the tail of the prior $\pi(\theta)$ is small enough such that the integral in the RHS of \eqref{eq:upper_bound} is finite. In general, this bound is tight whenever the support of the optimal distribution is a mass distribution at the origin. The following proposition implies that this bound is always tight as $\sigma / \sigma_\theta$ goes to infinity.
\begin{prop} \label{prop:asymp}
 For any prior $\pi(\theta)$ for which
\[
\int \frac{\pi(d\theta)}{\eta \left(\theta \right)} < \infty,
\]
we have 
\begin{align}
\label{eq:asymp}
\mathbb E [1/K(\theta)] = \frac{\pi}{2} + \left( \frac{\pi}{2} - 1\right) \left(\frac{\sigma_\theta}{\sigma} \right)^2 + o \left(\frac{1}{\sigma} \right)^3. 
\end{align}
as $\sigma_\theta / \sigma \rightarrow 0$. 
\end{prop}
\begin{proof}
It is enough to prove that  \eqref{eq:asymp} holds for the upper bound \eqref{eq:upper_bound}. First note that the condition in the proposition implies that all the moments of $\pi(\theta)$ exists, and in particular its second and third. The results follows by expanding the function $1/\eta(x)$ in a power series as
\[
1/\eta(x) = \frac{\pi}{2} + \left(\frac{\pi}{2}-1 \right) x^2 + o(x^3).
\]
\end{proof}

@@

We now consider the proof of Thm.~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$, we first note that without loss of generality, each message $M_i$ can be written in the form
\begin{equation}
\label{eq:general_messages}
M_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \mathbb R$ is a Lebesgue measurable set. Indeed, any measurable function $M(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = M^{-1}(1)$. Consider the conditional distribution $P({M^n|\theta})$ of $M^n$ given $\theta$. We have 
\begin{align}
P\left( M^n | \theta \right) & =  \prod_{i=1}^n P\left(M_i | \theta, M^{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(M_i =1 | \theta, M^{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$. We now prove the following Lemma:

Going back to \eqref{eq:adpt_lower_bound_proof:1}, it follows that the Fisher information of $M^n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(M^n) = \sum_{i=1}^n I_\theta (M_i|M^{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (M_i|M^{i-1})$ is the Fisher information of the distribution of $M_i$ given $M^{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (M_i|M^{i-1}) \leq 2/(\pi \sigma^2)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\mathbb E \left( \theta_n - \theta \right)^2 &  \geq \frac{1}{ \mathbb E I_\theta(M^n) + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (M_i | M^{i-1} ) + I_0} \\
& \geq \frac{1}{ 2n/(\pi \sigma^2) + I_0}.
\end{align*}

>>>>

@@

\subsection{Threshold Detection}
We assume that each message is of the form
\[
M_i = \sgn(t_i - X_i) = \begin{cases} 1 & X_i< t_i, \\
-1 & X_i > t_i,
\end{cases}  
\]
where $t_i\in\mathbb R$ is the \emph{threshold} of the $i$th detector applied by the encoder on it sample $X_i$. In order to characterize the asymptotic behavior of estimation from these sequence of messages, we consider the \emph{density} of $n$ threshold values defined as:
\[
\lambda_n([a,b]) = \frac{1}{n} \left| \mathcal T \cap [a,b] \right|,
\]
where $\mathcal T  = \left\{t_n, n \in \mathbb N \right\}$ is the collection of all threshold values. We further assume that $\lambda_n$ converges (weakly) to a probability measure $\lambda(t)$ on $\mathbb R$.  \par
In order to estimate $\theta$ from the messages $M^n$, we consider the maximum likelihood (ML) estimator. The log-likelihood function of $\theta$ is given by
\[
l(M^n  |\theta ) = \sigma  \sum_{i=1}^n \log \left( \Phi \left( M_i \frac{t_i - \theta}{\sigma} \right) \right). 
\]
Since $\Phi(x)$ is a log concave function, the log-likelihood function has a unique maximizer $\widehat{\theta}_n$ which is the ML estimator.  \\

@@

{\bf Worst case $\theta$}: in order to bound the local asymptotic minimax risk it is required to consider the desnity $\lambda$ that minimizes the worst case $\theta$ in the parameter space $\Theta$. We assume that $\Theta = [-b,b]$ for some $b>0$, and consider 
\begin{equation} \label{eq:variational}
\sup_{\lambda} \inf_{\theta \in [-b,b]}  {K(\theta)} = \int \eta\left( \frac{t-\theta}{\sigma} \right) \lambda (dt),
\end{equation}
where the suprimum is over all probability measure $\lambda$ on $\mathbb R$. Since \eqref{eq:variational} is concave in $\lambda$, this optimization problem can be solved using a convex program. 
As illustrated in Fig.~\ref{fig:cvx}, the threshold density that maximizes \eqref{eq:variational} is supported over a discrete set in the interval $[-b,b]$. Fig.~\ref{fig:cvx2} illustrates the asymptotic ML risk as a function of the length $2b$ of the parameter space $\theta$.

@@

As can be seen from Fig.~\ref{fig:opt_density}, for a normal prior with a small variance compared to the variance of the samples $\sigma^2$, the optimal density $\lambda$ is a mass distribution. As it turns out from the following proposition, the choice of  $\lambda$ to be a mass distribution leads to a general upper bound on the expected asymptotic risk, and not only on the minimal value of \eqref{eq:cvx_average}. 
\begin{prop}\label{prop:upper_bound}
Let $\theta_0 = \mathbb E \theta$. Then
\begin{equation}
\label{eq:upper_bound}
\mathbb E \frac{\sigma^2}{K(\theta)}  \leq \sigma^2 \int \frac{\pi(d\theta)}{\eta \left( \frac{\theta_0 - \theta}{\sigma} \right)}, 
\end{equation}
\end{prop}
\begin{proof}
Since the function $x \rightarrow 1/x$ is convex for positive values, Jensen's inequality applied to the RHS of \eqref{eq:cvx_average} leads to
\[
 \frac{\sigma^2}{K(\theta)} \leq \sigma^2 \int  \frac{ \lambda(dt)}{ \eta \left( \frac{t-\theta}{\sigma} \right)  }. 
\]
Therefore, the expected value of $\sigma^2/K(\theta)$ satisfies
\begin{align}
\mathbb E  \frac{\sigma^2}{K(\theta)}  \leq \sigma^2  \int \int \frac{\pi(d\theta) \lambda(dt) }{\eta \left( \frac{t - \theta}{\sigma} \right)}. \label{eq:upper_bound_proof}
\end{align}
The bound \eqref{eq:upper_bound} follows by using $\lambda = \delta_{\theta_0}$. 
\end{proof}
We note that since the function $1/\eta(x)$ is quasi-convex and symmetric, taking $\lambda = \lambda_{\theta_0}$ minimizes the RHS of \eqref{eq:upper_bound_proof} and leading to the tightest bound among the family bounds obtained after using Jensen's inequality.  \par
The bound \eqref{eq:upper_bound} is not trivial as long as the tail of the prior $\pi(\theta)$ is small enough such that the integral in the RHS of \eqref{eq:upper_bound} is finite. In general, this bound is tight whenever the support of the optimal distribution is a mass distribution at the origin. The following proposition implies that this bound is always tight as $\sigma / \sigma_\theta$ goes to infinity.
\begin{prop} \label{prop:asymp}
 For any prior $\pi(\theta)$ for which
\[
\int \frac{\pi(d\theta)}{\eta \left(\theta \right)} < \infty,
\]
we have 
\begin{align}
\label{eq:asymp}
\mathbb E [1/K(\theta)] = \frac{\pi}{2} + \left( \frac{\pi}{2} - 1\right) \left(\frac{\sigma_\theta}{\sigma} \right)^2 + o \left(\frac{1}{\sigma} \right)^3. 
\end{align}
as $\sigma_\theta / \sigma \rightarrow 0$. 
\end{prop}
\begin{proof}
It is enough to prove that  \eqref{eq:asymp} holds for the upper bound \eqref{eq:upper_bound}. First note that the condition in the proposition implies that all the moments of $\pi(\theta)$ exists, and in particular its second and third. The results follows by expanding the function $1/\eta(x)$ in a power series around zero as
\[
1/\eta(x) = \frac{\pi}{2} + \left(\frac{\pi}{2}-1 \right) x^2 + o(x^3).
\]
\end{proof}

@@

We now consider the proof of Thm.~\ref{thm:adpative_lower_bound}. In order to bound from above the Fisher information of any set of $n$ single-bit messages with respect to $\theta$, we first note that without loss of generality, each message $M_i$ can be written in the form
\begin{equation}
\label{eq:general_messages}
M_i = \begin{cases}
X_i \in A_i & 1, \\
X_i \notin A_i & -1,
\end{cases} 
\end{equation}
where $A_i \subset \mathbb R$ is a Lebesgue measurable set. Indeed, any measurable function $M(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = M^{-1}(1)$. Consider the conditional distribution $P({M^n|\theta})$ of $M^n$ given $\theta$. We have 
\begin{align}
P\left( M^n | \theta \right) & =  \prod_{i=1}^n P\left(M_i | \theta, M^{i-1} \right), \label{eq:adpt_lower_bound_proof:1}
\end{align}
where $P\left(M_i =1 | \theta, M^{i-1}  \right) = \mathbb P\left( X_i \in A_i\right)$. Going back to \eqref{eq:adpt_lower_bound_proof:1}, it follows that the Fisher information of $M^n$ with respect to $\theta$ is given by 
\begin{align}
I_\theta(M^n) = \sum_{i=1}^n I_\theta (M_i|M^{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (M_i|M^{i-1})$ is the Fisher information of the distribution of $M_i$ given $M^{i-1}$. From Lemma~\ref{lem:fisher_bound} it follows that $I_\theta (M_i|M^{i-1}) \leq 2/(\pi \sigma^2)$. The Van Trees inequality \cite{van2004detection, gill1995applications} now implies 
\begin{align*}
\mathbb E \left( \theta_n - \theta \right)^2 &  \geq \frac{1}{ \mathbb E I_\theta(M^n) + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (M_i | M^{i-1} ) + I_0} \\
& \geq \frac{1}{ 2n/(\pi \sigma^2) + I_0}.
\end{align*}
